---
title: "Reproducibility Report: Group B Choice 2"
output:
  html_document:
    toc: true
    toc_float: true
---

-------

For this exercise, please try to reproduce the results from Experiment 1 of the associated paper (Ko, Sadler & Galinsky, 2015). The PDF of the paper is included in the same folder as this Rmd file. 

#### Methods summary: 

A sense of power has often been tied to how we perceive each other's voice. Social hierarchy is embedded into the structure of society and provides a metric by which others relate to one another. In 1956, the Brunswik Lens Model was introduced to examine how vocal cues might influence hierarchy. In "The Sound of Power: Conveying and Detecting Hierarchical Rank Through Voice," Ko and colleagues investigated how manipulation of hierarchal rank within a situation might impact vocal acoustic cues. Using the Brunswik Model, six acoustic metrics were utilized (pitch mean & variability, loudness mean & variability, and resonance mean & variability) to isolate a potential contribution between individuals of different hierarchal rank. In the first experiment, Ko, Sadler & Galinsky examined the vocal acoustic cues of individuals before and after being assigned a hierarchal rank in a sample of 161 subjects (80 male). Each of the six hierarchy acoustic cues were analyzed with a 2 (high vs. low rank condition) x 2 (male vs. female) analysis of covariance, controlling for the baseline of the respective acoustic cue.

------

#### Target outcomes: 

Below is the specific result you will attempt to reproduce (quoted directly from the results section of Experiment 1):

> The impact of hierarchical rank on speakers’ acoustic cues. Each of the six hierarchy-based (i.e., postmanipulation) acoustic variables was submitted to a 2 (condition: high rank, low rank) × 2 (speaker’s sex: female, male) between-subjects analysis of covariance, controlling for the corresponding baseline acoustic variable. Table 4 presents the adjusted means by condition. Condition had a significant effect on pitch, pitch variability, and loudness variability. Speakers’ voices in the high-rank condition had higher pitch, **F(1, 156) = 4.48, p < .05**; were more variable in loudness, **F(1, 156) = 4.66, p < .05**; and were more monotone (i.e., less variable in pitch), **F(1, 156) = 4.73, p < .05**, compared with speakers’ voices in the low-rank condition (**all other Fs < 1**; see the Supplemental Material for additional analyses of covariance involving pitch and loudness). (from Ko et al., 2015, p. 6; emphasis added)

The adjusted means for these analyses are reported in Table 4 (Table4_AdjustedMeans.png, included in the same folder as this Rmd file).

------

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

## Step 1: Load packages

```{r}
library(tidyverse) # for data munging
library(knitr) # for kable table formating
library(haven) # import and export 'SPSS', 'Stata' and 'SAS' Files
library(readxl) # import excel files

# #optional packages:
library(psych)
 library(car) # for ANCOVA
 library(compute.es) # for ANCOVA
 library(lsmeans) # for ANCOVA
```

## Step 2: Load data

```{r}
# Just Experiment 1
d <-read_csv("data/S1_voice_level_Final.csv")

head(d)
```

## Step 3: Tidy data


Identified that "form" in DF means resonance, "intense" means loudness, and "pitch" is pitch. yet there are multiple cols for each of these characteristics, so I'm not sure which to use... Know that paper Adjusted means based on baseline, perhaps that is why there is 36 cols?? Does it just need to be cleaned? 

for example: form_smean, form,svar, and formrmean, and formrvar are different from fromant_rmeanMD, formant_rvarMD and Zform_smean, Zform_svar (are the latter standardized?)

form_smean form_svar form_rmean form_rvar
intense_smean intense_svar intense_rmean intense_rvar 
pitch_smean pitch_svar pitch_rmean pitch_rvar

pitch_smeanMD pitch_svarMD pitch_rmeanMD pitch_rvarMD 
intense_smeanMD intense_svarMD intense_rmeanMD intense_rvarMD 
formant_smeanMD formant_svarMD formant_rmeanMD formant_rvarMD 

Zpitch_smean Zpitch_svar Zpitch_rmean Zpitchrvar
Zintense_smean Zintense_svar Zintense_rmean Zintense_rvar
Zform_smean Zform_svar Zform_rmean Zform_rvar




```{r}



```

## Step 4: Run analysis

### Pre-processing

```{r}



```

### Descriptive statistics

In the paper, the adjusted means by condition are reported (see Table 4, or Table4_AdjustedMeans.png, included in the same folder as this Rmd file). Reproduce these values below:

```{r}
pitch_model <- aov(pitch_rmean ~ plev + pitch_smean, data = d)
summary(pitch_model)
 

```
### 3HR MARK 
### Inferential statistics

> The impact of hierarchical rank on speakers’ acoustic cues. Each of the six hierarchy-based (i.e., postmanipulation) acoustic variables was submitted to a 2 (condition: high rank, low rank) × 2 (speaker’s sex: female, male) between-subjects analysis of covariance, controlling for the corresponding baseline acoustic variable. [...] Condition had a significant effect on pitch, pitch variability, and loudness variability. Speakers’ voices in the high-rank condition had higher pitch, **F(1, 156) = 4.48, p < .05**; were more variable in loudness, **F(1, 156) = 4.66, p < .05**; and were more monotone (i.e., less variable in pitch), **F(1, 156) = 4.73, p < .05**, compared with speakers’ voices in the low-rank condition (**all other Fs < 1**; see the Supplemental Material for additional analyses of covariance involving pitch and loudness).

```{r}
# reproduce the above results here

```

# Step 5: Reflection

Were you able to reproduce the results you attempted to reproduce? If not, what part(s) were you unable to reproduce?

> ANSWER HERE

How difficult was it to reproduce your results? 


What aspects made it difficult? What aspects made it easy?

No coding manual, trying to reverse-engineer what the data actually mean, condition codes, why each collumn was included in the dataset. This consisted of hours of me scraatching muy head and flipping back and forth between this sheet and the paper. That was really frustrating. 